---
title: "Let's Code - Week 8"
author: "Adrien ROUX"
date: "February, 2017"
output: 
  html_document: 
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 1
---

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The topic introduced this week is **Data visualisation** which is a wide topic that covers : 

* 2d plots (histogram, points, lines),
* 3d plots (3d points, 3d lines,surface,...),
* geo maps,
* word clouds.

Further data visualisation are available, e.g. pie charts. If I had to present one more, it would be network representations. Unfortunaltely, I need more practice on this topic to be able to present you something.

***

# Word clouds

## A first example

We first read some text.

```{r}
text <- "Many years ago the great British explorer George Mallory, who 
was to die on Mount Everest, was asked why did he want to climb 
it. He said, \"Because it is there.\""
```

Now, we will perform a series of operations on the text data to simplify it.

```{r}
library(tm)
library(SnowballC)
library(wordcloud)

# 1. we need to create a corpus.
myCorpus <- Corpus(VectorSource(text)) #tm function

# 2. we convert the corpus to a plain text document.
myCorpus <- tm_map(myCorpus, PlainTextDocument)

# 3. we remove all punctuation and stopwords. 
# Stopwords are commonly used words in the English language such as I, me, my, etc. You can see the full list of stopwords using stopwords('english').

myCorpus <- tm_map(myCorpus, removePunctuation)
myCorpus <- tm_map(myCorpus, removeWords, stopwords('english'))

```

```{r}
# 4. we perform stemming. 
# This means that all the words are converted to their stem (Ex: learning -> learn, walked -> walk, etc.). This will ensure that different forms of the word are converted to the same form and plotted only once in the wordcloud.

myCorpus <- tm_map(myCorpus, 
                   stemDocument)
```

Now, we plot the wordcloud using a palette of colors instead of just one.

```{r}
library(RColorBrewer)
pal <- brewer.pal(6,"Dark2")
```

*** 

```{r, fig.align='center'}
 wordcloud(myCorpus, max.words = 100, 
          scale = c(4,0.5), random.order = FALSE,
          random.color = TRUE, colors = pal[-(1)],
          rot.per = 0.5)
```

## WordCloud2

Here is a somehow more advanced function to create word cloud with advanced features.

```{r, fig.align='center', warning=FALSE}
library(wordcloud2)
str(demoFreq)
wordcloud2(demoFreq)
head(demoFreq[order(demoFreq$freq, decreasing = T),],10)
```

Here is the same wordcloud with a shape of *star*.

```{r}
wordcloud2(demoFreq, size = 0.5, shape = 'star')
```

## Exercise

Load a text of your choice either stored locally or online and produce a first word cloud.

# 2d plots

3 different plotting systems/solutions:

* **standard method** using base package : plot (**ugly but simple** so fits your prupose if you quickly want to have an overview of your data in console mode),
* **lattice graphics** : emphasis on multivariate data (**ugly and complex** so I don't present it)
* **ggplot2** : *so pretty* but not so simple at the beginning. The best option to put graphs in a report. 

***

```{r, fig.align='center'}
x1 <- seq(from=-5,to=5,by=0.25)
y1 <- sapply(x1, function(x){x^2})
plot(x1,y1, type="point", main= "Example with plot function")

x2 <- seq(from=-5,to=5,by=0.5)
y2 <- sapply(x2, function(x){(x-1)^2})
lines(x2,y2, col="red")
```

## ggplot2

```{r, fig.align='center'}
library(ggplot2)
str(iris)
qplot(Sepal.Length, 
      Sepal.Width,
      data=iris, 
      color= Species) + facet_grid(.~Species)
```

# Geo maps

## leaflet

<https://rstudio.github.io/leaflet/>

```{r, fig.align='center'}
library(leaflet)

m <- leaflet() %>%
  setView(lng=2.35, lat = 48.85, zoom = 11) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=2.35, lat=48.85, popup="Center of Paris") %>%
  addMarkers(lng=2.23, lat=48.88, popup="La Defense") %>%
  addCircleMarkers(lng=2.3, lat=48.86, color = "#F00", radius= 20)
m  # Print the map
```

```{r}
# add some circles to a map
df = data.frame(Lat = 48.85 + rnorm(10,0,0.1), Long = 2.35 + rnorm(10,0,0.1))
leaflet(df) %>% 
  setView(lng=2.35, lat = 48.85, zoom = 10) %>%
  addTiles() %>%
  addCircles(lng = ~Long, lat = ~Lat, color = "#000")
```


# 3d plots

There is a lot of packages offering 3d visualisations. plotly is one of my favorites.
<https://plot.ly/r/>

```{r, warning=FALSE, message=FALSE}
library(plotly)
# economics : US economic time series
p <- plot_ly(economics, x = ~date, y = ~unemploy / pop)
p
```

```{r}
mtcars$am[which(mtcars$am == 0)] <- 'Automatic'
mtcars$am[which(mtcars$am == 1)] <- 'Manual'
mtcars$am <- as.factor(mtcars$am)

p <- plot_ly(mtcars, 
             x = ~wt, y = ~hp, z = ~qsec, 
             color = ~am, colors = c('#BF382A', '#0C4B8E')) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'Weight'),
                      yaxis = list(title = 'Gross horsepower'),
                      zaxis = list(title = '1/4 mile time')))
p
```

```{r}
plot_ly (
  # Using a dataframe:
  type = 'surface' , z = ~volcano )
```

```{r}
plot_ly(
  type = 'scatter3d', 
  x = c( 9, 8, 5, 1 ) , 
  y = c( 1, 2, 4, 8 ) , 
  z = c( 11, 8, 15, 3 ) ,
  mode = 'lines' )
```

# Conclusion

We reviewed various kinds of data visualisation achievable using R and extension packages. The legacy packages are now completely outdated. Most people use at least **ggplot2** and even **plotly** which allows interactive graphs using html.

Word clouds are really just for fun... 

# Assignement for next week

Practice at least one of each graph using any kind of data.
